Parameter Estimation for the CES Model

### 2013-07-30 CES.Rnw
bug fixes in running multiple CES models

<<>>=
dataUK <- loadData(countryAbbrev="UK")
modelUKQ_LM <- cesEst(data=dataUK, 
                      yName="iGDP", 
                      xNames=c("iCapStk", "iLabor", "iQ"), 
                      tName="iYear")
summary(modelUKQ_LM)
@

@@

<<>>=
modelUKQ_LM2 <- cesEst(data=dataUK, 
                      yName="iGDP", 
                      xNames=c("iLabor", "iQ", "iCapStk"), 
                      tName="iYear")
summary(modelUKQ_LM2)
@

@@

<<>>=
modelUKQ_LBFGSB <- cesEst(data=dataUK, 
                          yName="iGDP", 
                          xNames=c("iCapStk", "iLabor", "iQ"), 
                          tName="iYear",
                          method="L-BFGS-B")
summary(modelUKQ_LBFGSB)
@

@@

To assess what is happening, we can use the built-in grid search
capability of \texttt{cesEst}. Grid search is accessed by specifying
\texttt{rho} and/or \texttt{rho1} arguments and a range of values. 
Note that specifying a range of values for $\rho$ or $\rho_1$ that 
contains values less than $-1$ results in an error. 
The model with lowest $SSE$ will be returned from the grid search.

<<>>=
modelUKQ_grid <- cesEst(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor", "iQ"), 
                        tName="iYear",
                        method="L-BFGS-B",
                        rho=seq(from=5, to=30, by=5), 
                        rho1=seq(from=-0.5, to=1.5, by=0.5))
summary(modelUKQ_grid)
@

@@

<<>>=
modelUKQ_grid_detail <- cesEst(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor", "iQ"), 
                        tName="iYear",
                        method="L-BFGS-B",
                        rho=seq(from=100, to=120, length.out=20), 
                        rho1=seq(from=1, to=2, length.out=20))
summary(modelUKQ_grid_detail)
levelplot( RSS ~ rho1 + rho, data=extractRSSdata(modelUKQ_grid_detail), 
           region=TRUE, 
           col.regions=terrain.colors)
@

@@

<<>>=
modelUK <- cesEst(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor"), 
                        tName="iYear",
                        method="L-BFGS-B")
summary(modelUK)
@

@@

<<>>=
nestingOptions <- c(kle="(kl)e", lek="(le)k", ekl="(ek)l")
fitCES <- function(countryAbbrev, nest, energyType="Q", method="L-BFGS-B"){
  data <- loadData(countryAbbrev=countryAbbrev)
  data <- replaceColName(data, energyType, "iEToFit")
  if (nest == "(kl)e"){
    xNames <- c("iCapStk", "iLabor", "iEToFit")
  } else if (nest == "(le)k"){
    xNames <- c("iLabor", "iEToFit", "iCapStk")
  } else if (nest == "(ek)l"){
    xNames <- c("iEToFit", "iCapStk", "iLabor")
  } else {
    stop(paste("Unknown nesting option", nest, "in fitCES."))
  }
  model <- cesEst(data=data, 
                  yName="iGDP", 
                  xNames=xNames, 
                  tName="iYear",
                  method=method)
  res <- do(1) * coef( model ) 
  res <- transform( res, country=countryAbbrev, nest=nest, converge=model$convergence ) 
  return(res)
}

# Make all the models
kleModels <- ldply(countryAbbrevs, fitCES, nest="(kl)e")
lekModels <- ldply(countryAbbrevs, fitCES, nest="(le)k")
eklModels <- ldply(countryAbbrevs, fitCES, nest="(ek)l")
allModels <- rbind(kleModels, lekModels, eklModels)
# Ignore the leading .id column
allModels <- allModels[ , 2:10] 
# Sort rows based on country
allModels_ordered <- allModels[order(allModels$country) , ] 
# Reorder the columns to put the meta information first
allModels_ordered <- allModels_ordered[c("country", "nest", "converge", "gamma", "lambda", 
                                         "delta_1", "delta", "rho_1", "rho")]
# Create the LaTeX table
CESTable <- xtable(allModels_ordered, caption="CES coefficients for three nesting options.", 
                   label="tab:CES-coeffs-table",
                   table.placement="H")
@

<<CESResultsTable, results='asis'>>=
print(CESTable,
      caption.placement="top", 
      table.placement="H",
      include.rownames=FALSE,
      hline.after=c(-1,0,3,6,9,12,15,18,21,24,nrow(CESTable)))
@

>>>>

<<>>=
dataUK <- loadData(countryAbbrev="UK")
modelUKQ_LM <- cesEstPlus(data=dataUK, 
                      yName="iGDP", 
                      xNames=c("iCapStk", "iLabor", "iQ"), 
                      tName="iYear")
summary(modelUKQ_LM)
@

@@

<<>>=
modelUKQ_LM2 <- cesEstPlus(data=dataUK, 
                      yName="iGDP", 
                      xNames=c("iLabor", "iQ", "iCapStk"), 
                      tName="iYear")
summary(modelUKQ_LM2)
@

@@

<<>>=
modelUKQ_LBFGSB <- cesEstPlus(data=dataUK, 
                          yName="iGDP", 
                          xNames=c("iCapStk", "iLabor", "iQ"), 
                          tName="iYear",
                          method="L-BFGS-B")
summary(modelUKQ_LBFGSB)
@

@@

To assess what is happening, we can use the built-in grid search
capability of \texttt{cesEstPlus}. Grid search is accessed by specifying
\texttt{rho} and/or \texttt{rho1} arguments and a range of values. 
Note that specifying a range of values for $\rho$ or $\rho_1$ that 
contains values less than $-1$ results in an error. 
The model with lowest $SSE$ will be returned from the grid search.

<<>>=
modelUKQ_grid <- cesEstPlus(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor", "iQ"), 
                        tName="iYear",
                        method="L-BFGS-B",
                        rho=seq(from=5, to=30, by=5), 
                        rho1=seq(from=-0.5, to=1.5, by=0.5))
summary(modelUKQ_grid)
@

@@

<<>>=
modelUKQ_grid_detail <- cesEstPlus(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor", "iQ"), 
                        tName="iYear",
                        method="L-BFGS-B",
                        rho=seq(from=100, to=120, length.out=20), 
                        rho1=seq(from=1, to=2, length.out=20))
summary(modelUKQ_grid_detail)
levelplot( RSS ~ rho1 + rho, data=extractRSSdata(modelUKQ_grid_detail), 
           region=TRUE, 
           col.regions=terrain.colors)
@

@@

<<>>=
modelUK <- cesEstPlus(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor"), 
                        tName="iYear",
                        method="L-BFGS-B")
summary(modelUK)
@

@@

<<>>=
nestingOptions <- c(kle="(kl)e", lek="(le)k", ekl="(ek)l")
fitCES <- function(countryAbbrev, nest, energyType="Q", method=c("PORT","L-BFGS-B")){
  data <- loadData(countryAbbrev=countryAbbrev)
  data <- replaceColName(data, energyType, "iEToFit")
  if (nest == "(kl)e"){
    xNames <- c("iCapStk", "iLabor", "iEToFit")
  } else if (nest == "(le)k"){
    xNames <- c("iLabor", "iEToFit", "iCapStk")
  } else if (nest == "(ek)l"){
    xNames <- c("iEToFit", "iCapStk", "iLabor")
  } else {
    stop(paste("Unknown nesting option", nest, "in fitCES."))
  }
  model <- cesModel(data=data, energyType=energyType, 
                  # yName="iGDP", 
                  xNames=xNames, 
                  # tName="iYear",
                  method=method)
  res <- naturalCoef( model ) 
  res <- transform( res, nest=nest, country=countryAbbrev, nest=nest, converge=model$convergence ) 
  return(res)
}

# Make all the models
kleModels <- ldply(countryAbbrevs, fitCES, nest="(kl)e")
lekModels <- ldply(countryAbbrevs, fitCES, nest="(le)k")
eklModels <- ldply(countryAbbrevs, fitCES, nest="(ek)l")
allModels <- rbind(kleModels, lekModels, eklModels)
# Ignore the leading .id column
allModels <- subset( allModels, select= -.id )
allModels <- transform( allModels, max.sse = pmax(sse.PORT, sse.L.BFGS.B))
# Sort rows based on country
allModels_ordered <- allModels[order(allModels$country) , ] 
# Reorder the columns to put the meta information first
allModels_ordered <- allModels_ordered[c("country", "nest", "converge", "gamma", "lambda", 
                                         "delta_1", "delta", "rho_1", "rho", "method", "sse","max.sse")]
# Create the LaTeX table
CESTable <- xtable(allModels_ordered, caption="CES coefficients for three nesting options.", 
                   label="tab:CES-coeffs-table", digits=3,
                   table.placement="H")
@

<<CESResultsTable, results='asis'>>=
print(CESTable,
      caption.placement="top", 
      table.placement="H",
      include.rownames=FALSE,
      hline.after=c(-1,0,3,6,9,12,15,18,21,24,nrow(CESTable)),
      digits=3
      )
@
